---
title: "Distributed ARIMA"
abstract: ""
keywords: ""

course: Datenanalyse in Big Data (Prof. Dr. Buchwitz)
supervisor: Prof. Dr. Buchwitz
city: Meschede

# List of Authors
author:
- familyname: Hoheisel
  othernames: Jonas
  address: "MatNr: "
  qualifications: "Data Science (MA, 2. Semester)"
  email: hoheisel.jonas@fh-swf.de
- familyname: Henkenherm
  othernames: Kathrin
  address: "MatNr: 30362826"
  qualifications: "Data Science (MA, 2. Semester)"
  email: henkenherm.kathrin@fh-swf.de
- familyname: Katzenberger
  othernames: Ole
  address: "MatNr: "
  qualifications: "Data Science (MA, 2. Semester)"
  email: katzenberger.ole@fh-swf.de
- familyname: Krilov
  othernames: Vitali
  address: "MatNr: "
  qualifications: "Data Science (MA, 2. Semester)"
  email: krilov.vitali@fh-swf.de
- familyname: Stasenko
  othernames: Vladislav
  address: "MatNr: "
  qualifications: "Data Science (MA, 2. Semester)"
  email: stasenko.vladislav@fh-swf.de
# - familyname: Curie
#   othernames: Pierre
#   address: "MatNr: 87654321"
#   qualifications: "Data Science (MA, 2. Semester)"
#   email: curie.pierre@fh-swf.de

# Language Options
german: false # German Dummy Text
lang: en-gb   # Text Language: en-gb, en-us, de-de

# Indexes
toc: true     # Table of Contents
lot: false    # List of Tables
lof: false    # List of Figures

# Output Options
bibliography: references.bib
biblio-style: authoryear-comp
blind: false
cover: true
checklist: true
output:
  fhswf::seminarpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    keep_tex: no
    number_sections: yes
    citation_package: biblatex
knit: fhswf::render_seminarpaper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE, messages=FALSE, warning=FALSE, 
                      attr.source='.numberLines', singlespacing = TRUE)
fhswf::fhswf_hooks()

# Load Packages
library(fhswf)
library(ggplot2)
```


# Introduction

# ARIMA and Distributed Computing

# Map Reduce

## Reduce: Standard method

## Reduce: unconditional arithmetic mean

# Implementation with Google Cloud

# Forecast 

## Point Forecast

## Prediction Intervals

# Evaluation

## Accuracy: MASE and MSIS

## Performance: PySpark vs. Single Node

# Discussion

# Appendix

| Chapter      | Author     |
|--------------|-----------|
| Introduction |     |
| ARIMA and Distributed Computing      |  |
| Map Reduce      |  |
| Reduce: Standard method      |  |
| Reduce: unconditional arithmetic mean      |  |
| Implementation with Google Cloud      |  |
| Forecast      |  |
| Point Forecast      |  |
| Prediction Intervals      |  |
| Evaluation      |  |
| Accuracy: MASE and MSIS      |  |
| Performance: PySpark vs. Single Node      |  |
| Discussion    |  |



<!-- # Data -->
<!-- In a famous paper, @BC64 introduced a family of transformations \dots -->

<!-- ```{r histogram, fig.cap="Nice histogram.", message=FALSE, warning=FALSE} -->
<!-- qplot(exp(rnorm(200))) + theme_bw() -->
<!-- ``` -->

<!-- Consider the logNormal data plotted in Figure \ref{fig:histogram}. -->

<!-- # Methodology -->

<!-- Look a wild formula for $s^2$ appeared. -->

<!-- $$s^2 = \frac{1}{n-1}\sum_{i=1}^n (x_i-\bar{x})^2$$ -->

<!-- Longer Formulas can be spread across multiple lines manually using the `aligned` environment. -->

<!-- $$ -->
<!-- \begin{aligned} -->
<!-- \hat{y}_i = {} & \hat{\beta}_0 + \\ -->
<!--                & \hat{\beta}_1 \cdot x_1 +  \\ -->
<!--                & \hat{\beta}_2 \cdot x_2 + \hat{\beta}_3 \cdot x_3 + \hat{\epsilon} -->
<!-- \end{aligned} -->
<!-- $$ -->

<!-- # Analysis -->

<!-- ```{r cars} -->
<!-- knitr::kable(head(mtcars), booktabs=T, linesep="", caption="Some Cars.") -->
<!-- ``` -->

<!-- Table \ref{tab:cars} contains some values for cars \dots -->

<!-- This template supports execution of **R** and **Python** code. -->

<!-- ## R Code -->

<!-- ```{r, echo=T} -->
<!-- # seed random number generator and generate 3 random numbers -->
<!-- set.seed(1) -->
<!-- rnorm(n=3) -->
<!-- ``` -->

<!-- ## Python Code -->

<!-- ```{python, echo=T} -->
<!-- from random import seed -->
<!-- from random import random -->
<!-- # seed random number generator and generate 3 random numbers -->
<!-- seed(1) -->
<!-- print(random(), random(), random()) -->
<!-- ``` -->

<!-- # Citation -->

<!-- ## Using Citations -->

<!-- This templates comes with some frequently used references. -->

<!-- @StudienbuchStatistik -->

<!-- @FoliensatzStatistik -->

<!-- References can be cited in three different ways. -->

<!-- @Fahrmeir2016 -->

<!-- @Fahrmeir2016[p. 1058] -->

<!-- [@Fahrmeir2016, p. 1058] -->

<!-- ## Managing References -->

<!-- References can be added to the `references.bib` that comes with this template.  -->

<!-- # Conclusion -->

<!-- To switch the language to german, change the parameters in the `YAML` Header to the following values. -->

<!-- ```{r, echo=T, eval=F} -->
<!-- # Language Options -->
<!-- german: true -->
<!-- lang: de-de  -->
<!-- ``` -->

\newpage

# Technical Appendix {-}

```{r, echo = TRUE}
Sys.time()
sessionInfo()
```

